Working Notes

### 2019-02-26 NB_2r2c.Rmd
update

This page outlines Ver 2r.2c of NB Classifier and the process for building various components used in the project. Details of each  component are provided in these sub-pages.

For structure of the initial dataset see: https://adanieljohnson.github.io/default_website/datastructure.html

For import and post-import data validation see: https://adanieljohnson.github.io/default_website/dataimport.html

@@

#Initial Setup
Calling required libraries.
```{r libraries}
library(tidyverse)
library(tm)
library(tidytext)
library(e1071)
library(caret)
```

@@

##Graph Outputs
This code block generates two graphs comparing the sub-category frequencies obtained by hand-coding vs. predicted frequencies from the Naive Bayes classifier. There are separate comparisons for the Subject and Structure categories.

```{r}
blendSubject = ggplot()+
  geom_col(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`, group=1), color="blue", fill="cornflowerblue", alpha=0.5, width=0.4) +
  geom_col(data=proptallysum2, aes(`Bayes.Predicted.Subject`, `total`, group=1), color="brown", fill="indianred4", alpha=0.5, width=0.4,position = position_nudge(x=0.1))+
  xlab("Subject Groups")+
  ylab("Fraction of All Comments")+
  ggtitle("Comparison of Coded (blue) vs Predicted (red) Subject")

blendStructure = ggplot()+
  geom_col(data=proptallysum3, aes(`Hand.Coded.Structure`, `total`,group=1), color="springgreen4", fill="springgreen3", alpha=0.5, width=0.4) +
  geom_col(data=proptallysum4, aes(`Bayes.Predicted.Structure`, `total`, group=1), color="goldenrod3", fill="goldenrod2", alpha=0.5, width=0.4, position = position_nudge(x=0.1))+
  xlab("Structure Groups")+
  ylab("Fraction of All Comments")+
  ggtitle("Comparison of Coded (green) vs Predicted (orange) Structure")

print(blendSubject)
print(blendStructure)
```

>>>>

@@

#Initial Setup
Calling required libraries.
```{r libraries}
library(tidyverse)
library(tm)
library(tidytext)
library(e1071)
library(caret)
library(ggplot2)
```

@@

#Tabular and Graphical Comparisons of Coded and Predicted Categories
These code blocks compare the hand-coded versus NB-predicted frequencies. There are separate comparisons for the Subject and Structure categories, and both tabular and graphical comparisons.

```{r}
colnames(proptallysum1)[colnames(proptallysum1)=="Hand.Coded.Subject"] <- "Subject"
colnames(proptallysum1)[colnames(proptallysum1)=="total"] <- "Hand"
colnames(proptallysum2)[colnames(proptallysum2)=="Bayes.Predicted.Subject"] <- "Subject"
colnames(proptallysum2)[colnames(proptallysum2)=="total"] <- "Bayes" 
propmerge12<-merge.data.frame(proptallysum1,proptallysum2, all=TRUE)
propmerge12
```

```{r}
blendSubject = ggplot()+
  geom_col(data=proptallysum1, aes(`Subject`, `Hand`, group=1), color="blue", fill="cornflowerblue", alpha=0.5, width=0.4) +
  geom_col(data=proptallysum2, aes(`Subject`, `Bayes`, group=1), color="brown", fill="indianred4", alpha=0.5, width=0.4,position = position_nudge(x=0.12))+
  xlab("Subject Groups")+
  ylab("Fraction of All Comments")+
  ggtitle("Comment Subject Frequencies: Coded (blue) vs Predicted (red)")

print(blendSubject)
```

```{r}
colnames(proptallysum3)[colnames(proptallysum3)=="Hand.Coded.Structure"] <- "Structure"
colnames(proptallysum3)[colnames(proptallysum3)=="total"] <- "Hand"
colnames(proptallysum4)[colnames(proptallysum4)=="Bayes.Predicted.Structure"] <- "Structure"
colnames(proptallysum4)[colnames(proptallysum4)=="total"] <- "Bayes"
propmerge34<-merge.data.frame(proptallysum3,proptallysum4, all=TRUE)
propmerge34
```

```{r}
blendStructure = ggplot()+
  geom_col(data=proptallysum3, aes(`Structure`, `Hand`,group=1), color="springgreen4", fill="springgreen3", alpha=0.5, width=0.4) +
  geom_col(data=proptallysum4, aes(`Structure`, `Bayes`, group=1), color="goldenrod3", fill="goldenrod2", alpha=0.5, width=0.4, position = position_nudge(x=0.12))+
  xlab("Structure Groups")+
  ylab("Fraction of All Comments")+
  ggtitle("Comment Structure Frequencies: Coded (green) vs Predicted (orange)")

print(blendStructure)
```

####\  
***

### 2019-02-26 background.Rmd
update

---
title: "Background"
bibliography: library.bib
biblio-style: apalike
---

@@

Undergraduate introductory STEM courses often include technical writing training. Too often it is treated as an incidental skill that each student develops "along the way." At other times the instructional strategy is  haphazardly designed, implemented, and executed.

Adding to the learning challenge, faculty may not be the ones helping students develop writing skills. If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing. TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.

@@

*  Embody best practices from 40+ years of research on "writing in the disciplines;"
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from copy-correction towards larger writing issues; and
*  Can be implemented at scale by less experienced teaching assistants.

####\  
##Context & Prior Work
From prior studies (*see the Literature Reveiw for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing. Ideally:

@@

Our bins-based report grading protocol reinforces this feedback strategy.

This holistic, coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy. We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable. What is needed is a more direct measure of TA effort.

@@

We next used a subset of 110 randomly selected comments to build an initial qualitative codebook for classifying comments based on subject and structure. This initial codebook was tested and revised using blocks of 1100 additional randomly selected comments until the categories and criteria stabilized. Using the final codebook, we then sorted all 11,000 comments into separate categories (see PAGE_LINK for complete codebook.)

@@

>>>>

---
title: "Background of This Project"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

@@

Undergraduate introductory STEM courses often include technical writing training. Too often it is treated as an incidental skill that each student develops "along the way." At other times the instructional strategy is  haphazardly designed, implemented, and executed [@holder].

Adding to the learning challenge, faculty may not be the ones helping students develop writing skills. If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing [@holder]. TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way [@holder].

@@

*  Embody best practices from 40+ years of research on "writing in the disciplines";
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from copy-correction towards larger writing issues; and
*  Can be implemented at scale by less experienced teaching assistants.

####\  
##Context & Prior Work
From prior studies (*see the Literature Reveiw for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead [@holder]. To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing. Ideally:

@@

Our bins-based report grading protocol [@Nilson] reinforces this feedback strategy.

This holistic, coaching-oriented approach is well-supported by prior research [@holder], but we are uncertain how accurately TAs implement the recommended strategy. We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable [@holder]. What is needed is a more direct measure of TA effort.

@@

We next used a subset of 110 randomly selected comments to build an initial qualitative codebook [@holder] for classifying comments based on subject and structure. This initial codebook was tested and revised using blocks of 1100 additional randomly selected comments until the categories and criteria stabilized. Using the final codebook, we then sorted all 11,000 comments into separate categories (see PAGE_LINK for complete codebook.)

@@

####\  
***

### 2019-02-26 build.Rmd
updates
update

---
title: "The Build"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Overview
This page links to sub-pages that describe how various components used in the project were built. Follow the links for details.

Extracting TAs' comments: https://adanieljohnson.github.io/default_website/docxcommentextraction.html

Structure of development dataset; data import & validation: https://adanieljohnson.github.io/default_website/dataimport.html

Current Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html

####\

#Describe an Element

####\

#Describe an Element

####\

>>>>

---
title: "Project Build-Out"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

This page links to sub-pages describing construction of the various components used in the overall project. Progress on each component is in parentheses.

**Extracting TAs' Comments** *(Early coding stage)*  
Protocol used to extract TA comments from student lab reports.  
https://adanieljohnson.github.io/default_website/docxcommentextraction.html

**Structure of Development Dataset; Data Import & Validation** *(Completed)*  
Organization of the initial dataset, and how data were imported in R and validated.  
https://adanieljohnson.github.io/default_website/dataimport.html

**Codebook for Classifying TAs' Comments** *(Completed)*  
Describes the process for developing the codes used to classify comments.  
https://adanieljohnson.github.io/default_website/codebook.html

**Initial Exploration of the Dataset** *(Partly completed)*  
What is the distribution of words used in the TA comments?  
https://adanieljohnson.github.io/default_website/initial_exploration.html

**Current Version of the Naive Bayes Comment Classifier** *(Completed)*  
Code for the most recent version of the optimized NB classifier.   
https://adanieljohnson.github.io/default_website/NB_2r2c.html

**Embedding the Naive Bayes Classifier in an Ensemble Model** *(Planning stage)*  
The NB method alone does not classify TA comments with sufficient accuracy for the "Structure" category. This page outlines a vertically stacked "ensemble" classifier that uses pattern matching and other pre-processing steps to improve NB accuracy.
https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

####\  
***

### 2019-02-26 index.Rmd
update

#### \   
##Work Plan
The work plan for this project is:

1.  Compile an initial dataset for testing, and de-identify both TAs and students by replacing names with anonymous IDs.
2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
3.  Use basic text analysis methods to explore comment data structure, n-gram frequencies, etc., and identify potentially useful elements for feature engineering.
4.  Write a baseline workflow using Naive Bayes for supervised text classification.
5.  Assign TA comments to categories using the NB classifier, then establish baseline accuracy values using the optimized NB classifier by comparing outcomes of automated classification to hand-coded results.
6.  Using features (from Step #3) and comparison data from NB classification protocol (from Step #5) as guides, further optimize the NB classifier method. 
7.  When accuracy stabilizes, apply the optimized NB classifier to the original FULL comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier to create a vertical Ensemble Classifier.
10.  Validate the Ensemble Classifier using a subset of ~2000 new hand-coded TA comments extracted from a non-overlapping set of student lab reports. Compare automated results for new data set to hand-coded categories.

Looking beyond the current project, I will be using the NB-based Ensemble Classifier as the control method for evaluating other text classifier strategies.

@@

#### \

>>>>

#### \   
##Work Plan
The work outline for this project is:

1.  Compile an initial dataset of hand-coded TA comments on reports for testing. Data will be de-identified by replacing both TAs' and students' names with anonymous IDs.
2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
3.  Use basic text analysis methods to explore comment data structure, n-gram frequencies, etc., and identify potentially useful elements for feature engineering.
4.  Write a workflow using Naive Bayes for supervised text classification, and test various permutations of analysis parameters to establish a baseline classifier.
5.  Assign TA comments to categories using the baseline NB classifier, and establish baseline accuracy values by comparing outcomes of automated classification to hand-coded results.
6.  Using features (from Step #3) and comparison data from NB classification protocol (from Step #5) as guides, further optimize the NB classifier method until predicted accuracy stabilizes.
7.  Apply the optimized NB classifier to the original FULL comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
9.  Combine the rules/pattern-based pre-screening process with the optimized Naive Bayes classifier to create a vertical Ensemble Classifier.
10.  Re-validate the Ensemble Classifier using a subset of ~2000 new hand-coded TA comments extracted from student lab reports for a different semester than the initial test dataset. Compare automated results for new data set to hand-coded categories.

Looking beyond the current project, I will use the vertical Ensemble Classifier two ways:

*  To monitor TA comment patterns in the ongoing NSF-funded project.
*  As the control method for evaluating other potential text classifier strategies.

@@

#### \   
***

### 2019-02-26 litreview.Rmd
update

#Overview
The supporting literature for this project is broad and convoluted. Rather than try to compress it to a single page, I have provided links out to sub-pages that describe various background topics. Follow the links for details.

Fundamental Principles and Practices of Technical Writing Training:  https://adanieljohnson.github.io/default_website/pagename.html

Primer on Text Classification Strategies:  https://adanieljohnson.github.io/default_website/Text_classifier_models.html

####\

>>>>

This project brings together three disparate challenges:

*  The challenges of teaching technical writing in the STEM disciplines
*  Needs and barriers for training future faculty to teach technical writing
*  Automated text analysis and classification

Follow the links below (or from top menu) to sub-pages that explore each of these challenges in more detail.

**Challenges of Teaching Technical Writing**  
More than 50% of colleges and universities use some form of Writing in the Disciplines, Write to Learn, and Writing Across the Curriculum (WID/WTL/WAC) as part of their general education programs. There are well-established WAC/WID teaching principles based on >40 years of evidence and experience, but much of this work focuses on non-technical styles of writing. We know much less about teaching technical writing specifically. What do we know, and what can we bring from the WAC/WID community? What do we NOT know?  
https://adanieljohnson.github.io/default_website/techwriting.html

**Challenges of Training Future Faculty**  
A major barrier to routine writing in STEM is limited instructor awareness of evidence-based training models. Many college STEM teachers have never been introduced to proven writing training strategies or trained to give effective, context-specific feedback. How does this impact instruction more generally?  
https://adanieljohnson.github.io/default_website/futurefaculty.html

**Challenges of Applying Automated Text Analysis to Student Learning**  
Interest in technology-assisted automated feedback on writing began in the mid-1960s. Systems have progressed from basic spelling and grammar checkers to automated essay scoring (AES) systems that provide feedback on the content of open-response writing. Can automated text analysis support student learning?  
https://adanieljohnson.github.io/default_website/aacr.html

**Text Classification 1 - Feature Selection and Engineering**  
What is the general strategy and steps for solving a text classification problem? What features are important? 
https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html

**Text Classification 2 - Topic Categorization and Modeling**  
Establishing meaningful text classification categories depends on the goals of the project or the question being asked. Are relevant categories known already? If not, how are they established?  
https://adanieljohnson.github.io/default_website/Topic_categorization.html

**Text Classification 3 - Selecting an Appropriate Classifier Method**  
There are three general classes of text classification strategies: pattern matching, algorithmic, and neural networks. Each is described briefly.  
https://adanieljohnson.github.io/default_website/Choosing_classifier.html

**Text Classification 4 - Evaluating and Improving the Classifier**  
If there are many ways to classify texts, how do we know which one is most accurate? How can we improve the classification process? How do we make comparisons and evaluate the different steps in a text classification problem?  
https://adanieljohnson.github.io/default_website/Improving_classifier.html

####\  
***

### 2019-02-26 nextsteps.Rmd
update

Qualitative coding of TA comments is informative but unsustainable. Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook, individual coders will require training to achieve any accuracy. Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.

####\

>>>>

DRAFT PLACEHOLDER PAGE

Lorem ipsum dolor sit amet...

####\  
***

### 2019-02-26 notes.rmd
update

##General/Site-Wide:
  Embedding the bibliographic items  
    Writing the bibtext  
    Embedding using [@ BaseR] method.

##Index:
*  Decide on bottom page site map strategy

@@

##Lit review:  
*  Df  
    +  Df

@@

##Findings:
*  How to graph the different comparisons of NB values from different sources on same plot? 
    +  Add ggplot (data=) value.
    +  Better to combine data into one dataset though.

@@

#Etc.:
*  Using a Bibliography in this format  
*  Adding automatic updates to top dates? 
    +  ```date: "`r format(Sys.time(), '%b %d, %Y')`" ```

*  Fix GitHub auto-upload

@@

####\

>>>>

##General/Site-Wide:
  Embedding the bibliographic items  
    Writing the bibtext  
    Embedding using @baseR method, or [@holder].  
  Modify the icons:

  Set menu so landing page is an "Overview"
  Append this code to end of all pages
```
      #### \   
      ***
```

##Index:
*  Decide on bottom page site map strategy, and attach

@@

##Lit review:  
*  Finalize what content pages will be
    +  Df

@@

##Findings:
*  How to graph the different comparisons of NB values from different sources on same plot? 
    +  Add ggplot (data=) value. SOLVED 
    +  Better to combine data into one dataset though. (REQUIRED for proper legend formats)

@@

#Etc.:
*  Adding the Bibliography in this format
    +  Need to add the new citation entries to library.bib
*  Adding automatic updates to top dates? 
    +  ```date: "`r format(Sys.time(), '%b %d, %Y')`" ```

#Miscellaneous:
*  Fix GitHub link on laptop to auto-upload without passwords again

@@

####\  
888
